---
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "man/figures/")
```

# moire <img src="man/figures/logo.svg" align="right" alt="" height="139" />

`moire` is a package implementing an MCMC based approach to estimating complexity of infection and population allele frequencies from polyallelic genomics data. 

## Installation
`moire` requires the Gnu Scientific Library (GSL). 

```{bash, eval = FALSE}
# Linux
sudo apt-get install libgsl-dev
```

```{bash, eval = FALSE}
# MacOS
brew install gsl
```

For Windows, follow instructions found [here](https://stackoverflow.com/questions/26939683/linking-gsl-library-to-rcppgsl-on-windows-machine) 

After installing GSL, install moire as follows
```{r, eval = FALSE}
# Install development version from Github
remotes::install_github("m-murphy/moire")
```

## Usage
moire supports loading data from either a long format `data.frame` using `load_long_form_data()` or from a wide format `data.frame` using `load_delimited_data()`.

```{r, eval = FALSE}
df <- read.csv("your_data.csv")
data <- load_long_form_data(df)

# With data in appropriate format, run MCMC as follows
mcmc_results <- moire::run_mcmc(data$data, data$sample_ids, data$loci, data$is_missing)
```
