---
output: github_document
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.path = "man/figures/")
```

# moire <img src="man/figures/logo.svg" align="right" alt="" height="139" />

moire is a package implementing an MCMC based approach to estimating complexity of infection and population allele frequencies from polyallelic genomics data. Details of the method may be found [here](https://github.com/m-murphy/moire/raw/master/refs/thesis.pdf).

## Installation

```{r, eval = FALSE}
# Install development version from Github
remotes::install_github("m-murphy/moire")
```

## Usage
moire supports loading data from either a long format `data.frame` using `load_long_form_data()` or from a wide format `data.frame` using `load_delimited_data()`.

```{r, eval = FALSE}
df <- read.csv("your_data.csv")
data <- load_long_form_data(df)

# With data in appropriate format, run MCMC as follows
mcmc_results <- moire::run_mcmc(data$data, data$sample_ids, data$loci, data$is_missing)
```
