---
title: "MCMC Demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MCMC Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(moire)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tidyr)
```

To demonstrate the usage of the package, we can simulate some genotyping data according to our model.

```{r simulate_data, eval=FALSE, include=TRUE}
mean_moi <- 4
num_loci <- 10
num_biological_samples <- 100
epsilon_pos <- .03
epsilon_neg <- .15

# Generate the number of alleles at each locus
allele_counts = rnbinom(num_loci, size = 15, mu = 10) + 2

# We'll use flat alpha vectors for our draws from the Dirichlet
locus_freq_alphas = lapply(allele_counts, function(allele) { rep(1, allele) })

simulated_data <- moire::simulate_data(mean_moi, locus_freq_alphas, num_biological_samples, 
                                       epsilon_pos, epsilon_neg)
```

```{r load_simulated_data, eval=TRUE, include=FALSE}
mean_moi <- 4
num_loci <- 10
num_biological_samples <- 100
epsilon_pos <- .03
epsilon_neg <- .15
simulated_data <- moire::simulated_data
```

Now that we have our data, let's go ahead and run the MCMC 

```{r run_mcmc, eval=FALSE, include=TRUE}
burnin = 1e3
num_samples = 1e3
res <- run_mcmc(simulated_data$data, burnin = burnin, samples = num_samples, eps_pos_0 = .05, eps_neg_0 = .1,
                eps_pos_var = .1, eps_neg_var = .1, max_coi = 20, importance_sampling_depth = 100,
                eps_pos_alpha = 1, eps_pos_beta = 1, eps_neg_alpha = 1, eps_neg_beta = 1)
```

```{r run_mcmc_load, include = FALSE}
res <- moire::mcmc_results
burnin <- length(res$loglike_burnin)
num_samples <- length(res$loglike_sample)
```

After running the MCMC, we can access the draws from the posterior distribution and analyze.

```{r analyze_mcmc}
# Estimate the COI for each sample
est_cois = matrix(unlist(res$m_store), ncol = num_samples)
est_coi_025 = apply(est_cois, 1, function(x) { quantile(x, .025) })
est_coi_50 = apply(est_cois, 1, function(x) { quantile(x, .5) })
est_coi_975 = apply(est_cois, 1, function(x) { quantile(x, .975) })
est_coi_mean = apply(est_cois, 1, mean)


# Calculate naive COI for comparison
naive_cois <- moire::calculate_naive_coi(simulated_data$data)
offset_naive_cois <- moire::calculate_naive_coi_offset(simulated_data$data, 2)
coi_data <- data.frame(est_coi_025, est_coi_50, est_coi_975, est_coi_mean,
                       naive_coi = naive_cois, offset_naive_coi = offset_naive_cois)


# Calculate allele frequencies from MCMC model
est_allele_freqs = matrix(unlist(res$p_store), ncol = num_samples)
est_allele_freq_025 = apply(est_allele_freqs, 1, function(x) { quantile(x, .025)})
est_allele_freq_50 = apply(est_allele_freqs, 1, function(x) { quantile(x, .5)})
est_allele_freq_975 = apply(est_allele_freqs, 1, function(x) { quantile(x, .975)})
est_allele_freq_mean = apply(est_allele_freqs, 1, mean)


# Calculate Heterozygosity for each locus from model
he_matrix <- matrix(unlist(lapply(res$p_store, function(run) {
  sapply(run, moire::calculate_he)
  })), ncol = num_samples)

est_he_025 <- apply(he_matrix, 1, function(x) { quantile(x, .025)})
est_he_50 <- apply(he_matrix, 1, function(x) { quantile(x, .5)})
est_he_975 <- apply(he_matrix, 1, function(x) { quantile(x, .975)})
est_he_mean <- apply(he_matrix, 2, mean)
est_he_mean_quantiles <- quantile(est_he_mean, c(.025, .5, .975))
```

Here we demonstrate the difference in distribution of the estimate of MOI using our approach vs the naive method of taking the maximum number of alleles observed or the second highest number of alleles observed

```{r plot_analysis}
coi_estimates <- coi_data %>%
  select(est_coi_50, naive_coi, offset_naive_coi) %>%
  gather(key = "estimator", value = "estimate") %>%
  mutate(estimator_pretty = case_when(
    estimator == 'est_coi_50' ~ "Mean Posterior Est.",
    estimator == 'naive_coi' ~ "Naive Estimate",
    estimator == 'offset_naive_coi' ~ "Offset Naive Estimate"
  )) %>%
  transform(estimator_pretty = factor(estimator_pretty, levels = c("Mean Posterior Est.", 
                                                                   "Offset Naive Estimate", 
                                                                   "Naive Estimate")))

ggplot(coi_estimates, aes(x = estimate, stat(count))) +
  facet_wrap(~ estimator_pretty) +
  geom_bar() +
  xlab("Multiplicity of Infection") +
  ylab("Total Count") +
  theme_classic(base_size = 16)

```

