---
title: "MCMC Demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MCMC Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup_options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
# 
do.call(knitr::read_chunk, list(path = system.file(c("data-raw", "mcmc_results.R"), package="moire")))
# knitr::read_chunk("../data-raw/mcmc_results.R")
```

```{r setup, message=FALSE}
library(moire)
library(magrittr)
library(ggplot2)
library(dplyr)
library(tidyr)
```

To demonstrate the usage of the package, we can simulate some genotyping data according to our model.
```{r load_settings, eval=TRUE, include=TRUE}

```

```{r simulate_data, eval=FALSE, include=TRUE}

```

```{r load_simulated_data, eval=TRUE, include=FALSE}
simulated_data <- moire::simulated_data
```

Now that we have our data, let's go ahead and run the MCMC. 

```{r run_mcmc, eval=FALSE, include=TRUE}
```

```{r run_mcmc_load, eval = TRUE, include = FALSE}
mcmc_results <- moire::mcmc_results
```
After running the MCMC, we can access the draws from the posterior distribution and analyze.

```{r analyze_mcmc}
# Estimate the COI for each sample
coi_summary <- moire::summarize_coi(mcmc_results)

# We can also summarize statistics about the allele frequency distribution
he_summary <- moire::summarize_he(mcmc_results)
allele_freq_summary <- moire::summarize_allele_freqs(mcmc_results)

# Let's append the true values for use later
coi_data <- data.frame(coi_summary,
  true_coi = simulated_data$sample_cois
  )

he_data <- data.frame(
  he_summary,
  true_he = sapply(
    moire::calculate_naive_allele_frequencies(simulated_data$true_genotypes),
    function(x) moire::calculate_he(x)
  ),
  naive_he = sapply(
    moire::calculate_naive_allele_frequencies(simulated_data$data),
    function(x) moire::calculate_he(x)
  )
)

allele_freq_data <- data.frame(
  allele_freq_summary,
  naive_allele_frequency = unlist(
    moire::calculate_naive_allele_frequencies(simulated_data$data)
  ),
  true_allele_frequency = unlist(
    moire::calculate_naive_allele_frequencies(simulated_data$true_genotypes)
  )
)
```

Here we demonstrate the difference in distribution of the estimate of MOI using our approach vs the naive method of taking the maximum number of alleles observed or the second highest number of alleles observed

```{r plot_distribution}
coi_estimates <- coi_data %>%
  dplyr::select(post_coi_med, naive_coi, offset_naive_coi, true_coi) %>%
  tidyr::pivot_longer(tidyr::everything(),
                      names_to = "estimator",
                      values_to = "estimate") %>%
  dplyr::mutate(estimator_pretty = case_when(
    estimator == "post_coi_med" ~ "Median Posterior Est.",
    estimator == "naive_coi" ~ "Naive Estimate",
    estimator == "offset_naive_coi" ~ "Offset Naive Estimate",
    estimator == "true_coi" ~ "True COI")) %>%
  transform(estimator_pretty = factor(estimator_pretty, levels = c(
    "Median Posterior Est.",
    "Offset Naive Estimate",
    "Naive Estimate",
    "True COI"
  )))

ggplot(coi_estimates, aes(x = estimate, stat(count))) +
  facet_wrap(~estimator_pretty, nrow = 1) +
  geom_bar() +
  xlab("Complexity of Infection") +
  ylab("Total Count") +
  theme_classic(base_size = 12) +
  theme(axis.text.x = element_text(angle = 90))
```

Since we simulated data, we know the true complexity of infection and allele frequencies and can compare to our estimates.

```{r plot_coi_concordance}
ggplot(coi_data, aes(x = true_coi, y = post_coi_mean)) +
  geom_boxplot(aes(group = true_coi), color = "black", alpha = .1) +
  geom_jitter(shape = 1, alpha = .5, color = "black", width = .1, height = .1) +
  geom_jitter(aes(y = naive_coi), color = "red",
              alpha = .25, width = .1, height = .1) +
  geom_jitter(aes(y = offset_naive_coi), color = "green",
              alpha = .25, width = .1, height = .1) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Mean Posterior COI") +
  xlab("True COI") +
  theme_classic(base_size = 12) +
  expand_limits(x = 0, y = 0) +
  ggtitle("Estimates of COI vs Truth")
```
As expected, the naive estimator (in red) and the offset naive estimator (in green) are pretty heavily biased and. Complexity of infection is well estimated, but there's some difficulty in the upper range.

```{r plot_allele_freq_concordance}
ggplot(allele_freq_data) +
  geom_errorbar(aes(
    y = post_allele_freqs_mean,
    x = true_allele_frequency,
    ymax = post_allele_freqs_upper,
    ymin = post_allele_freqs_lower)) +
  geom_point(aes(y = post_allele_freqs_mean, x = true_allele_frequency)) +
  geom_point(aes(y = naive_allele_frequency, x = true_allele_frequency), color = "red", alpha = .3) +
  geom_abline(slope = 1, intercept = 0) +
  ylab("Mean Posterior Freqeuncy") +
  xlab("True Frequency") +
  theme_classic(base_size = 12) +
  expand_limits(x = 0, y = 0) +
  ggtitle("Estimates of Allele Frequencies vs Truth")
```
Again, naive estimation of allele frequencies (red) is heavily biased and lacks any estimates of uncertainty. Allele frequencies are very well estimated by MOIRE (black), recovering values from the entire range along with measures of uncertainty at the 95% credible interval. Since we have access to the full distribution of sample allele frequencies, we can also look at distributions of other summary statistics that might be of interest. For example, let's take a look at expected heterozygosity.
```{r plot_he_concordance}
ggplot(he_data, aes(x = true_he)) +
  geom_errorbar(aes(y = post_stat_mean, ymax = post_stat_upper, ymin = post_stat_lower)) +
  geom_point(aes(y = post_stat_mean), color = "blue") +
  geom_point(aes(y = naive_he), color = "green") +
  geom_abline(slope = 1, intercept = 0) +
  xlab("True Heterozygosity") +
  ylab("Mean Posterior Heterozygosity") +
  theme_classic()
```
Naive estimation (green) results in a severe upward bias, especially at lower levels of expected heterozygosity. MOIRE maintains high accuracy (blue), including measures of uncertainty at the 95% credible interval. 
